Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit

### 2024-06-30 optim.qmd
conclusions drafted

The primary goal of the study reported here is to investigate the performance of JSO in PP optimisation for the guided tour. It is of interest to assess how quickly and closely the optimiser reaches a gobal optima, for various PP indexes that may have differing complexities. To observe the performance of JSO with different types of PP indexes, metrics are introduced to capture specific properties of the index including squintability (based on @barnett1981interpreting's squint angle) and smoothness as introduced in @laa_using_2020. Here we mathematically define metrics for squintability and smoothness, which is new for the field.  A series of simulation experiments using specially structured datasets and various PP indexes are conducted to assess the behaviour of JSO under different settings and its sensitivity to hyper-parameter choices.  The relationship between the JSO performance, hyper-parameter tuning and properties of PP
indexes is analysed to provide guidance for practitioners using projection pursuit, and developers of new indexes.

@@

JSO provides a new swarm-based search strategy for optimising
projection pursuit problems. This section describes the setups for two simulations
to demonstrate JSO's performance. The first simulation compares JSO's 
performance with an exist optimiser, Creeping Random Search (CRS), and explores how JSO behaves
with different hyper-parameters combinations. The second simulation
studies the effect of optimisation properties (smoothness and squintability),
along with JSO hyper-parameters, on the success rate of the
optimisation.

@@

In this simulation, 50 jellyfishes (`n_jellies`) and
a maximum of 50 tries (`max_tries`) are used. The full data contains 50
simulations, indexed by the variable `sim`, with the associated `seed`s. 
Recorded in columns `basis`, `index_val`, `tries`, `loop` and
`time`, are the sampled projection basis, index value,
iteration ID, jellyfish ID, and computational time
respectively.

@@

In the second experiment, fifty simulations are conducted at each
hyper-parameter combination below to analyse its effects on the JSO success
rate: number of jellyfish: 20, 50, and 100, and  maximum number of 
tries: 50 and 100). The success 
rate is defined as the proportion of
simulations that achieves a final index value within 0.05 of the best
index value found among all 50 simulations. @fig-success-rate illustrates 
the calculation based on the projection found by the above simulation. Here
the optimal index value is
`r sim_sine_6d_spline_best |> pull(index_val) |> max() |> round(3) -> aa; aa`.
Out of the 50 simulations,
`r sim_sine_6d_spline_best |> filter(index_val >= aa - 0.05) |> nrow() -> bb; bb`
achieved an index value within \[`r aa - 0.05`, `r aa`\], resulting
in a success rate of
`r sim_sine_6d_spline_best |> mutate(a = index_val >= aa - 0.05) |> pull(a) -> cc; round(sum(cc)/length(cc), 3)`.

@@

## Factors affecting JSO success rate: optimisation properties and jellyfish hyper-parameters {#sec-app-2}

@@

The results from the two simulations described in @sec-sim-deets are analysed based on the final projections across two optimisers (JSO and CRS) and the success rate across JSO hyper-parameters. Smoothness and squintability are calculated across a collection of pipe-finding and sine-wave finding problems to construct the relationship between success rate, JSO hyper-parameters, and optimisation properties.

@@

The optimisation properties, including smoothness and squintability,
offer numerical metrics to characterise the complexity of
projection pursuit optimisation problems. @tbl-smoothness-squintability presents 
the parameters estimated from the
Gaussian process (variance, range, smooth, and nugget) and the scaled
logistic function ($\theta_1$ to $\theta_4$) for calculating both metrics for each case considered in the simulation. The column
"smooth" is used as the smoothness metric and the column "squint" is
calculated as equation @eq-squintability-parametric as the squintability metric.  @tbl-mod-output presents the results of fitting a generalised linear model with a binomial family and a logit link
function to the data in @tbl-mod-data, where all the three components (success rate, JS hyper-parameters, and optimisation properties) are combined. The model suggests that JSO success rate is positively
associated with the two hyper-parameters, as well as with the
optimisation properties: smoothness and squintability. Specifically,
using 10 more jellyfishes and 10 more tries increases the odd ratio of
success by `r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 2)`%
and `r round(100 * (exp(broom::tidy(mod1)$estimate[7])-1), 2)`%,
respectively. However, being flagged with long runtime and an increase
of data dimension reduce the success rate by
`r round(exp(broom::tidy(mod1)$estimate[5]) * 100, 2)`% and
`r round(exp(broom::tidy(mod1)$estimate[4]) * 100, 2)`%, respectively.
The variable `squintability` and `dimension` are significant, suggesting
their importance relative to JSO hyper-parameters in the optimisation
success. In defining a projection pursuit problem, factors such as the shape-to-find, the index function used, and the data dimension, determine the optimisation properties like smoothness and squintability. These metrics can then be compared across different problems to understand their relative complexity. The results suggest that squintability has a more significant influence than smoothness on the success rate of JSO. Once the characteristics of the projection pursuit problem are fully understood, increasing JSO hyper-parameters can enhance the search effectiveness, however, it is also important to consider the resulting increase in computational complexity.

@@

```{r}
#| label: tbl-mod-output
#| tbl-cap: Model estimates of proportion of jellyfish success on optimisation properties and jellyfish hyper-parameters from the generalised linear model with a binomial family and a logit link function. The variable smoothness, squintability, number of jellyfish and maximum number of tries are positively associated with JSO success rate while data dimension and being flagged as long runtime are negatively associated with the success rate. The variable squintability and dimension are significant, suggesting their importance relative to jellyfish hyper-parameters in the optimisation success.
broom::tidy(mod1) |> 
  mutate(term = c("Intercept", "Smoothness",
   "Squintability", "Dimension (d)", 
   "Long time", "N. jellyfish", 
   "Max. tries")) |>
  kable(digits = 2, format = "latex", 
    linesep = "", booktabs = T) |> 
  column_spec(c(1,5), border_right = TRUE) |> 
  column_spec(1, border_left = TRUE) |> 
  kable_styling(font_size = 10) 
```

@@

- Summarise the result of the simulation: JSO immense improvement on current methods, ...
- JSO for a guided tour, what changes are needed? Do we follow one tentacle, do we perform off-line and then pull tentacles out to watch, how does it integrate with ferrn

@@

  - Smoothness and squintability can be calculated and compared across different problems to understand their relative complexity. Once a projection pursuit problem is fully characterised (by the shape-to-find, the index function used, and the data dimension), increasing JSO hyper-parameters can enhance the search effectiveness, however, the resulting increase in computational complexity will also need to be considered. - modelling

>>>>

The primary goal of the study reported here is to investigate the performance of JSO in PP optimisation for the guided tour. It is of interest to assess how quickly and closely the optimiser reaches a gobal optima, for various PP indexes that may have differing complexities. To observe the performance of JSO with different types of PP indexes, metrics are introduced to capture specific properties of the index including squintability (based on @barnett1981interpreting's squint angle) and smoothness as introduced in @laa_using_2020. Here we mathematically define metrics for squintability and smoothness, which is new for the field.  A series of simulation experiments using specially structured datasets and various PP indexes are conducted to assess the behaviour of JSO under different settings and its sensitivity to hyper-parameter choices (number of jellies and maximum number of tries).  The relationship between the JSO performance, hyper-parameter tuning and properties of PP indexes is analysed to provide guidance for practitioners using projection pursuit, and developers of new indexes.

@@

The JSO performance is compared with the existing Creeping Random Search (CRS) currently used in guided tour, and explores how JSO behaves with different hyper-parameters combinations. A second simulation
studies the effect of index properties (smoothness and squintability), along with JSO hyper-parameters, on the success rate of the JSO performance.

@@

In this simulation, 50 jellyfishes (`n_jellies`) and
a maximum of 50 tries (`max_tries`) are used. The full data contains 50 simulations, indexed by the variable `sim`, with the associated `seed`s. Recorded in columns `basis`, `index_val`, `tries`, `loop` and `time`, are the sampled projection basis, index value, iteration ID, jellyfish ID, and computational time, respectively.

@@

In the second experiment, fifty simulations are conducted at each hyper-parameter combination below to analyse its effects on the JSO success rate: number of jellyfish: 20, 50, and 100, and  maximum number of  tries: 50 and 100). The success rate is defined as the proportion of simulations that achieves a final index value within 0.05 of the best index value found among all 50 simulations. @fig-success-rate illustrates 
the calculation based on the projection found by the above simulation. Here, the optimal index value is
`r sim_sine_6d_spline_best |> pull(index_val) |> max() |> round(3) -> aa; aa`. Out of the 50 simulations,
`r sim_sine_6d_spline_best |> filter(index_val >= aa - 0.05) |> nrow() -> bb; bb` achieved an index value within \[`r aa - 0.05`, `r aa`\], resulting in a success rate of `r sim_sine_6d_spline_best |> mutate(a = index_val >= aa - 0.05) |> pull(a) -> cc; round(sum(cc)/length(cc), 3)`.

@@

## Factors affecting JSO success rate: index properties and jellyfish hyper-parameters {#sec-app-2}

@@

The results from the two simulations described in @sec-sim-deets are analysed based on the final projections across two optimisers (JSO and CRS) and the success rate across JSO hyper-parameters. Smoothness and squintability are calculated across a collection of pipe-finding and sine-wave finding problems to construct the relationship between success rate, JSO hyper-parameters, and index properties.

@@

The index properties, including smoothness and squintability, offer numerical metrics to characterise the complexity of projection pursuit optimisation problems. @tbl-smoothness-squintability presents 
the parameters estimated from the Gaussian process (variance, range, smooth, and nugget) and the scaled
logistic function ($\theta_1$ to $\theta_4$) for calculating both metrics for each case considered in the simulation. The column
"smooth" is used as the smoothness metric and the column "squint" is
calculated as equation @eq-squintability-parametric as the squintability metric.  @tbl-mod-output presents the results of fitting a generalised linear model with a binomial family and a logit link
function to the data in @tbl-mod-data, where all the three components (success rate, JS hyper-parameters, and index properties) are combined. The model suggests that JSO success rate is positively associated with the two hyper-parameters, as well as with the index properties: smoothness and squintability. Specifically,
using 10 more jellyfishes and 10 more tries increases the odd ratio of success by `r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 2)`%
and `r round(100 * (exp(broom::tidy(mod1)$estimate[7])-1), 2)`%,
respectively. However, being flagged with long runtime and an increase of data dimension reduce the success rate by `r round(exp(broom::tidy(mod1)$estimate[5]) * 100, 2)`% and `r round(exp(broom::tidy(mod1)$estimate[4]) * 100, 2)`%, respectively. The variable `squintability` and `dimension` are significant, suggesting their importance relative to JSO hyper-parameters in the optimisation success. In defining a projection pursuit problem, factors such as the shape-to-find, the index function used, and the data dimension, determine the properties like smoothness and squintability. These metrics can then be compared across different problems to understand their relative complexity. The results suggest that squintability has a more significant influence than smoothness on the success rate of JSO. Once the characteristics of the projection pursuit problem are fully understood, increasing JSO hyper-parameters can enhance the search effectiveness, however, it is also important to consider the resulting increase in computational complexity.

@@

```{r}
#| label: tbl-mod-output
#| tbl-cap: Model estimates of proportion of jellyfish success on index properties and jellyfish hyper-parameters from the generalised linear model with a binomial family and a logit link function. The variable smoothness, squintability, number of jellyfish and maximum number of tries are positively associated with JSO success rate while data dimension and being flagged as long runtime are negatively associated with the success rate. The variable squintability and dimension are significant, suggesting their importance relative to jellyfish hyper-parameters in the optimisation success.
broom::tidy(mod1) |> 
  mutate(term = c("Intercept", "Smoothness",
   "Squintability", "Dimension (d)", 
   "Long time", "N. jellyfish", 
   "Max. tries")) |>
  kable(digits = 2, format = "latex", 
    linesep = "", booktabs = T) |> 
  column_spec(c(1,5), border_right = TRUE) |> 
  column_spec(1, border_left = TRUE) |> 
  kable_styling(font_size = 10) 
```

@@

This paper has presented new metrics to mathematically define desirable features of PP indexes, squintability and smoothness, and used these to assess the performance of the new jellyfish search optimiser. The metrics will be generally useful for characterising PP indexes, and halp with developing nex indexes.

In the comparison of the JSO against the currently used CRS, as expected the JSO vastly out-performs CRS, and provides a high probability of finding the global optima. The JSO obtains the maximum more cleanly, with a slightly higher index value, and plot of the projected data showing the structure more clearly.

The JSO performance is affected by the hyper-parameters, with a higher chance of reaching the global optima when more jellies are used and the maximum number of tries is increased. However, it comes at a computational cost, as expected. The performance declines if the projection dimension increases and if the PP index has low squintability. The higher the squintability the better chance the JSO can find the optima. However, interestingly smoothness does not affect the JSO performance.

The new JSO is integrated with the current implementation of the projection pursuit guided tour in the `tourr` package, and can be further examined using PP optimisation diagnostics in the `ferrn` package. Using the JSO is a little different than the current CRS, because it runs many paths. The recommended approach is to conduct the optimisation off-line, and then use the guided tour to follow selected jellies. The tools for this are available in the `tourr` package, too.

<!--
- Summarise the result of the simulation: JSO immense improvement on current methods, ...
- JSO for a guided tour, what changes are needed? Do we follow one tentacle, do we perform off-line and then pull tentacles out to watch, how does it integrate with ferrn

@@

  - Smoothness and squintability can be calculated and compared across different problems to understand their relative complexity. Once a projection pursuit problem is fully characterised (by the shape-to-find, the index function used, and the data dimension), increasing JSO hyper-parameters can enhance the search effectiveness, however, the resulting increase in computational complexity will also need to be considered. - modelling
-->

### 2024-06-30 optim.tex
conclusions drafted

The primary goal of the study reported here is to investigate the
performance of JSO in PP optimisation for the guided tour. It is of
interest to assess how quickly and closely the optimiser reaches a gobal
optima, for various PP indexes that may have differing complexities. To
observe the performance of JSO with different types of PP indexes,
metrics are introduced to capture specific properties of the index
including squintability (based on \citet{barnett1981interpreting}'s
squint angle) and smoothness as introduced in \citet{laa_using_2020}.
Here we mathematically define metrics for squintability and smoothness,
which is new for the field. A series of simulation experiments using
specially structured datasets and various PP indexes are conducted to
assess the behaviour of JSO under different settings and its sensitivity
to hyper-parameter choices. The relationship between the JSO
performance, hyper-parameter tuning and properties of PP indexes is
analysed to provide guidance for practitioners using projection pursuit,
and developers of new indexes.

@@

\begin{tcolorbox}[enhanced jigsaw, left=2mm, colback=white, opacityback=0, leftrule=.75mm, bottomtitle=1mm, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toptitle=1mm, colframe=quarto-callout-note-color-frame, rightrule=.15mm, coltitle=black, breakable, bottomrule=.15mm, titlerule=0mm, toprule=.15mm, opacitybacktitle=0.6]

@@

In this simulation, 50 jellyfishes (\texttt{n\_jellies}) and a maximum
of 50 tries (\texttt{max\_tries}) are used. The full data contains 50
simulations, indexed by the variable \texttt{sim}, with the associated
\texttt{seed}s. Recorded in columns \texttt{basis}, \texttt{index\_val},
\texttt{tries}, \texttt{loop} and \texttt{time}, are the sampled
projection basis, index value, iteration ID, jellyfish ID, and
computational time respectively.

In the second experiment, fifty simulations are conducted at each
hyper-parameter combination below to analyse its effects on the JSO
success rate: number of jellyfish: 20, 50, and 100, and maximum number
of tries: 50 and 100). The success rate is defined as the proportion of
simulations that achieves a final index value within 0.05 of the best
index value found among all 50 simulations.
Figure~\ref{fig-success-rate} illustrates the calculation based on the
projection found by the above simulation. Here the optimal index value
is 0.999. Out of the 50 simulations, 44 achieved an index value within
{[}0.949, 0.999{]}, resulting in a success rate of 0.88.

@@

\subsection{Factors affecting JSO success rate: optimisation properties
and jellyfish hyper-parameters}\label{sec-app-2}

@@

The results from the two simulations described in
Section~\ref{sec-sim-deets} are analysed based on the final projections
across two optimisers (JSO and CRS) and the success rate across JSO
hyper-parameters. Smoothness and squintability are calculated across a
collection of pipe-finding and sine-wave finding problems to construct
the relationship between success rate, JSO hyper-parameters, and
optimisation properties.

@@

\subsection{Factors affecting JSO success rate: optimisation properties
and jellyfish hyper-parameters}\label{sec-app-2}

The optimisation properties, including smoothness and squintability,
offer numerical metrics to characterise the complexity of projection
pursuit optimisation problems. Table~\ref{tbl-smoothness-squintability}
presents the parameters estimated from the Gaussian process (variance,
range, smooth, and nugget) and the scaled logistic function
(\(\theta_1\) to \(\theta_4\)) for calculating both metrics for each
case considered in the simulation. The column ``smooth'' is used as the
smoothness metric and the column ``squint'' is calculated as equation
\ref{eq-squintability-parametric} as the squintability metric.
Table~\ref{tbl-mod-output} presents the results of fitting a generalised
linear model with a binomial family and a logit link function to the
data in Table~\ref{tbl-mod-data}, where all the three components
(success rate, JS hyper-parameters, and optimisation properties) are
combined. The model suggests that JSO success rate is positively
associated with the two hyper-parameters, as well as with the
optimisation properties: smoothness and squintability. Specifically,
using 10 more jellyfishes and 10 more tries increases the odd ratio of
success by 24.11\% and 11.93\%, respectively. However, being flagged
with long runtime and an increase of data dimension reduce the success
rate by 41.72\% and 53.36\%, respectively. The variable
\texttt{squintability} and \texttt{dimension} are significant,
suggesting their importance relative to JSO hyper-parameters in the
optimisation success. In defining a projection pursuit problem, factors
such as the shape-to-find, the index function used, and the data
dimension, determine the optimisation properties like smoothness and
squintability. These metrics can then be compared across different
problems to understand their relative complexity. The results suggest
that squintability has a more significant influence than smoothness on
the success rate of JSO. Once the characteristics of the projection
pursuit problem are fully understood, increasing JSO hyper-parameters
can enhance the search effectiveness, however, it is also important to
consider the resulting increase in computational complexity.

@@

\caption{\label{tbl-mod-output}Model estimates of proportion of
jellyfish success on optimisation properties and jellyfish
hyper-parameters from the generalised linear model with a binomial
family and a logit link function. The variable smoothness,
squintability, number of jellyfish and maximum number of tries are
positively associated with JSO success rate while data dimension and
being flagged as long runtime are negatively associated with the success
rate. The variable squintability and dimension are significant,
suggesting their importance relative to jellyfish hyper-parameters in
the optimisation success.}

@@

\begin{itemize}
\tightlist
\item
  Summarise the result of the simulation: JSO immense improvement on
  current methods, \ldots{}
\item
  JSO for a guided tour, what changes are needed? Do we follow one
  tentacle, do we perform off-line and then pull tentacles out to watch,
  how does it integrate with ferrn
\end{itemize}

Main findings:

\begin{itemize}
\item
  JSO tends to obtian a clearer projection and higher index value when
  searching higher dimension spaces than CRS (and other current
  optimisers) - Figure 4
\item
  Increasing JS hyper-parameter enhances its performance, but the
  computational time required for the optimisation can quickly become
  intensive, especially when evaluating the index function (such as
  scagnostic indexes) multiple times across numerous iterations. -
  Figure 5
\item
  Smoothness and squintability can be calculated and compared across
  different problems to understand their relative complexity. Once a
  projection pursuit problem is fully characterised (by the
  shape-to-find, the index function used, and the data dimension),
  increasing JSO hyper-parameters can enhance the search effectiveness,
  however, the resulting increase in computational complexity will also
  need to be considered. - modelling
\end{itemize}

>>>>

The primary goal of the study reported here is to investigate the
performance of JSO in PP optimisation for the guided tour. It is of
interest to assess how quickly and closely the optimiser reaches a gobal
optima, for various PP indexes that may have differing complexities. To
observe the performance of JSO with different types of PP indexes,
metrics are introduced to capture specific properties of the index
including squintability (based on \citet{barnett1981interpreting}'s
squint angle) and smoothness as introduced in \citet{laa_using_2020}.
Here we mathematically define metrics for squintability and smoothness,
which is new for the field. A series of simulation experiments using
specially structured datasets and various PP indexes are conducted to
assess the behaviour of JSO under different settings and its sensitivity
to hyper-parameter choices (number of jellies and maximum number of
tries). The relationship between the JSO performance, hyper-parameter
tuning and properties of PP indexes is analysed to provide guidance for
practitioners using projection pursuit, and developers of new indexes.

@@

\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, bottomtitle=1mm, bottomrule=.15mm, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, leftrule=.75mm, coltitle=black, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, opacityback=0, colback=white, left=2mm, toprule=.15mm, arc=.35mm, colframe=quarto-callout-note-color-frame]

@@

In this simulation, 50 jellyfishes (\texttt{n\_jellies}) and a maximum
of 50 tries (\texttt{max\_tries}) are used. The full data contains 50
simulations, indexed by the variable \texttt{sim}, with the associated
\texttt{seed}s. Recorded in columns \texttt{basis}, \texttt{index\_val},
\texttt{tries}, \texttt{loop} and \texttt{time}, are the sampled
projection basis, index value, iteration ID, jellyfish ID, and
computational time, respectively.

In the second experiment, fifty simulations are conducted at each
hyper-parameter combination below to analyse its effects on the JSO
success rate: number of jellyfish: 20, 50, and 100, and maximum number
of tries: 50 and 100). The success rate is defined as the proportion of
simulations that achieves a final index value within 0.05 of the best
index value found among all 50 simulations.
Figure~\ref{fig-success-rate} illustrates the calculation based on the
projection found by the above simulation. Here, the optimal index value
is 0.999. Out of the 50 simulations, 44 achieved an index value within
{[}0.949, 0.999{]}, resulting in a success rate of 0.88.

@@

\subsection{Factors affecting JSO success rate: index properties and
jellyfish hyper-parameters}\label{sec-app-2}

@@

The results from the two simulations described in
Section~\ref{sec-sim-deets} are analysed based on the final projections
across two optimisers (JSO and CRS) and the success rate across JSO
hyper-parameters. Smoothness and squintability are calculated across a
collection of pipe-finding and sine-wave finding problems to construct
the relationship between success rate, JSO hyper-parameters, and index
properties.

@@

\subsection{Factors affecting JSO success rate: index properties and
jellyfish hyper-parameters}\label{sec-app-2}

The index properties, including smoothness and squintability, offer
numerical metrics to characterise the complexity of projection pursuit
optimisation problems. Table~\ref{tbl-smoothness-squintability} presents
the parameters estimated from the Gaussian process (variance, range,
smooth, and nugget) and the scaled logistic function (\(\theta_1\) to
\(\theta_4\)) for calculating both metrics for each case considered in
the simulation. The column ``smooth'' is used as the smoothness metric
and the column ``squint'' is calculated as equation
\ref{eq-squintability-parametric} as the squintability metric.
Table~\ref{tbl-mod-output} presents the results of fitting a generalised
linear model with a binomial family and a logit link function to the
data in Table~\ref{tbl-mod-data}, where all the three components
(success rate, JS hyper-parameters, and index properties) are combined.
The model suggests that JSO success rate is positively associated with
the two hyper-parameters, as well as with the index properties:
smoothness and squintability. Specifically, using 10 more jellyfishes
and 10 more tries increases the odd ratio of success by 24.11\% and
11.93\%, respectively. However, being flagged with long runtime and an
increase of data dimension reduce the success rate by 41.72\% and
53.36\%, respectively. The variable \texttt{squintability} and
\texttt{dimension} are significant, suggesting their importance relative
to JSO hyper-parameters in the optimisation success. In defining a
projection pursuit problem, factors such as the shape-to-find, the index
function used, and the data dimension, determine the properties like
smoothness and squintability. These metrics can then be compared across
different problems to understand their relative complexity. The results
suggest that squintability has a more significant influence than
smoothness on the success rate of JSO. Once the characteristics of the
projection pursuit problem are fully understood, increasing JSO
hyper-parameters can enhance the search effectiveness, however, it is
also important to consider the resulting increase in computational
complexity.

@@

\caption{\label{tbl-mod-output}Model estimates of proportion of
jellyfish success on index properties and jellyfish hyper-parameters
from the generalised linear model with a binomial family and a logit
link function. The variable smoothness, squintability, number of
jellyfish and maximum number of tries are positively associated with JSO
success rate while data dimension and being flagged as long runtime are
negatively associated with the success rate. The variable squintability
and dimension are significant, suggesting their importance relative to
jellyfish hyper-parameters in the optimisation success.}

@@

This paper has presented new metrics to mathematically define desirable
features of PP indexes, squintability and smoothness, and used these to
assess the performance of the new jellyfish search optimiser. The
metrics will be generally useful for characterising PP indexes, and halp
with developing nex indexes.

In the comparison of the JSO against the currently used CRS, as expected
the JSO vastly out-performs CRS, and provides a high probability of
finding the global optima. The JSO obtains the maximum more cleanly,
with a slightly higher index value, and plot of the projected data
showing the structure more clearly.

The JSO performance is affected by the hyper-parameters, with a higher
chance of reaching the global optima when more jellies are used and the
maximum number of tries is increased. However, it comes at a
computational cost, as expected. The performance declines if the
projection dimension increases and if the PP index has low
squintability. The higher the squintability the better chance the JSO
can find the optima. However, interestingly smoothness does not affect
the JSO performance.

The new JSO is integrated with the current implementation of the
projection pursuit guided tour in the \texttt{tourr} package, and can be
further examined using PP optimisation diagnostics in the \texttt{ferrn}
package. Using the JSO is a little different than the current CRS,
because it runs many paths. The recommended approach is to conduct the
optimisation off-line, and then use the guided tour to follow selected
jellies. The tools for this are available in the \texttt{tourr} package,
too.
